Compositional equivalences based on Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2019-04-24 2016-Weak-FH-Bisim/WeakBisim.tex
work on proof of new lemma 7

@@

\begin{lemma}[Open transition composition]\label{lem-compose}

\end{lemma}
\TODO{important comment added:}
Note that this does not mean that any two pNets can be composed and produce an open 
transition. Indeed, the predicate $\Pred\land\Pred\,'\land \alpha_Q=\beta_{j_0}$ is not 
be satisfiable if the action of $\alpha_Q$ cannot be matched with $\beta_{j_0}$.
Note also that $\beta_{j_0}$ is now only used as an intermediate term inside formulas: it 
does not appear neither as global action nor as an action of a hole.
\begin{small}

@@

@@

\begin{lemma}[Weak open transition composition]\label{lem-Weakcompose} Suppose that we have a weak open automaton such that the WOTs cannot observe silent actions (see Definition \ref{def:Non-ObsTau}).
	Suppose $j_0\in J$ and $\gamma_{j_0}=[\beta_0..\beta_n]$ and additionally:\\[-2ex]
\begin{mathpar}
%\[
P\models{\openrule
	{
		\gamma_j^{j\in J}, 
		\Pred,  
		\Post}
	{\ostate{s_i^{i\in L}} \OTWeakarrow {\alpha}
		\ostate{s_i'^{\, i\in L}}}
}\\
\text{~~and for all $p\in[0..n]$~~}
Q\models{\openrule
	{
		\gamma_{p j}^{j\in J_p},
		 \Pred_p,  
		\Post_p}
	{\ostate{s_{p i}^{i\in L_Q}} \OTWeakarrow {\alpha_p}
		\ostate{s_{(p+1) i}^{\, i\in L_Q}}}
}%\]
\end{mathpar}
Let 
\begin{mathpar}
J_Q=\displaystyle{\bigcup_{p=0}^nJ_{p}}
\qquad
\Pred\,'=\Pred\land \bigwedge_{p=0}^{n}(\alpha_p=\beta_p\land \Pred_p)\bigotimes_{i=0}^{p-1}\Post_i
\\
\forall j\in J_p,\,\gamma_j= \displaystyle{\mybigdotcup_{p=0}^n \gamma_{p j}}\text{\quad and\quad}
\Post\,'=\Post\uplus\bigotimes_{p=0}^{n}
		\Post_p 
\end{mathpar}
Then, we have\\[-2ex]
	\[ P[Q]_{j_0}  
	\models
	{\openrule
		{
			\gamma_j^{j\in (J\setminus\{j_0\}) \uplus J_Q}, 
			\Pred\,',  \Post\,'
			 }
		{\ostate{s_i^{i\in L\uplus L_Q}} \OTWeakarrow {\alpha}
			\ostate{s_i'^{\, i\in L\uplus L_Q}}}
	}
	\]
\end{lemma}
\TODO{Proof starts from here}

@@

Note that if $(\beta^i_p)_{j_0}=\tau$ because of Definition \ref{def:Non-ObsTau}, $P$ necessarily makes a $\tau$ open transition and remains in the same state, ($s_p = s_{p+1}$). Thus without loss of generality, we can bypass such an open transition and obtain another decomposition of the WOT without the open transition that requires ${\beta^i_{p{j_0}}}=\tau$. We can thus  suppose that for all $p$ and $i$ we have ${\beta^i_{p{j_0}}}\neq\tau$.

Moreover, we have for all $p\in[0..n]$:
\begin{mathpar}
Q\models{\openrule
	{
		\gamma_{p j}^{j\in J_p},
		 \Pred_p,  
		\Post_p}
	{\ostate{s_{p i}^{i\in L_Q}} \OTWeakarrow {\alpha_p}
		\ostate{s_{(p+1) i}^{\, i\in L_Q}}}
}		
\end{mathpar}
%In the same way,  by Lemma \ref{lem-rel-OT-WOT} we can decompose $\WOT_p$ into  a series of $m'+n'$ tau open transitions and a single observable open transition:
%\[Q\models{\openrule
%			{
%				(\beta^2_j)^{j\in J\cap\Holes(Q)}, {{\Pred'^{2}}},  
%				{{\Post'^{2}}}}
%				{\ostate{s_{p i}^{i\in L_Q}} \OTarrow {\alpha_p}
%				\ostate{s_{(p+1) i}^{i\in L_Q}}}
%				}%\]
%\]

We proceed by iteration on the list of  OTs of $P$ identified above, these OTs are of the form:

$P\!\models		
\openrule
    {
       {(\beta^i_{p j})^{j\in J}},\Pred_p,\Post_p   }
         {\ostate{s^{i\in L}_{pi}} \OTarrow {\alpha} \ostate{s^{i\in L}_{(p+1)i}}}$

@@

By Lemma \ref{lem-rel-OT-WOT} from cases $(1)$, $2$ and $(3)$ we obtain: 
\[P[Q]_{j_0}\!\models		
\openrule
    {
       \beta_p^{(p\in J\setminus\{j_0\}) \uplus J_Q}, {\Pred_p}
		\land {\Pred_p}\!\!'\land \alpha_p'=\beta_{j_0}$, ${\Post_p}\!\!''={\Post_p}\uplus 
		{\Post_p}\!\!'}
         {\ostate{s^{i\in L\uplus L_Q}_{pi}} \OTWeakarrow {\alpha} \ostate{s^{i\in L \uplus L_Q}_{(p+1)i}}}\]

 \item Else, $j_0\not\in J$. $t$ does not move  and $Q$ does not move, i.e. we have: $\ostate{(s_{ti})^{i\in L_Q}} = \ostate{({s}_{(t+1)i})^{i\in L_Q}}$....
\end{itemize}

>>>>

\TODO{voisr si on redefinit les bigotimes pour prendre des maps et retourner des maps au lieu des substitutions (et on englobe dans une subst a la fin)}

@@

\begin{lemma}[Open transition composition]\label{lem-compose} 
	Suppose $j_0\in J$ and:\\[-2ex]
\begin{mathpar}
%\[
P\models{\openrule
	{
		\beta_j^{j\in J}, 
		\Pred,  
		\Post}
	{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
		\ostate{s_i'^{\, i\in L}}}
}%\]
\text{~~and~~}
%\[
Q\models{\openrule
	{
		\beta_j^{j\in J_Q},
		 \Pred\,',  
		\Post\,'}
	{\ostate{s_i^{i\in L_Q}} \OTarrow {\alpha_Q}
		\ostate{s_i'^{\, i\in L_Q}}}
}%\]
\end{mathpar}
Then, we have\\[-2ex]
	\[ P[Q]_{j_0}  
	\models
	{\openrule
		{
			\beta_j^{(j\in J\setminus\{j_0\}) \uplus J_Q}, 
			\Pred\land\Pred\,'\land \alpha_Q=\beta_{j_0},  
			\Post\uplus \Post\,'}
		{\ostate{s_i^{i\in L\uplus L_Q}} \OTarrow {\alpha}
			\ostate{s_i'^{\, i\in L\uplus L_Q}}}
	}
	\]
\end{lemma}
\TODO{important comment added:}
Note that this does not mean that any two pNets can be composed and produce an open 
transition. Indeed, the predicate $\Pred\land\Pred\,'\land \alpha_Q=\beta_{j_0}$ is not 
be satisfiable if the action of $\alpha_Q$ cannot be matched with $\beta_{j_0}$.
Note also that $\beta_{j_0}$ is now only used as an intermediate term inside formulas: it 
does not appear neither as global action nor as an action of a hole.
\begin{small}

@@

Note that we also have the following lemma (trivial):

\begin{lemma}[Open transition composition]\label{lem-compose-2} ~\\	Suppose $j_0\not\in J$ and $L_Q=\Leaves({Q})$:\\[-2ex]
\begin{mathpar}
%\[
P\models{\openrule
	{
		\beta_j^{j\in J}, 
		\Pred,  
		\Post}
	{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
		\ostate{s_i'^{\, i\in L}}}
}%\]
\end{mathpar}
Then, we have\\[-2ex]
	\[ P[Q]_{j_0}  
	\models
	{\openrule
		{
			\beta_j^{j\in J}, 
			\Pred,  
			\Post}
		{\ostate{s_i^{i\in L}\uplus s_i^{i\in L_Q}} \OTarrow {\alpha}
			\ostate{s_i'^{\, i\in L}\uplus s_i^{i\in L_Q}}}}
	\]
\end{lemma}

@@

\begin{lemma}[Weak open transition composition]\label{lem-Weakcompose} Suppose that we have a weak open automaton such that the WOTs cannot observe silent actions (see Definition \ref{def:Non-ObsTau}).
	Suppose $j_0\in J$  and additionally:\\[-2ex]
\begin{mathpar}
%\[
P\models{\openrule
	{
		\beta_j^{j\in J}, 
		\Pred,  
		\Post}
	{\ostate{s_i^{i\in L}} \OTarrow {\alpha}
		\ostate{s_i'^{\, i\in L}}}
}\\
\text{~~and~~}
Q\models{\openrule
	{
		\set{\gamma},
		 \Pred_Q,  
		\Post_Q}
	{\ostate{s_{i}^{i\in L_Q}} \OTWeakarrow {\alpha_Q}
		\ostate{{s'}_{i}^{\, i\in L_Q}}}
}%\]
\end{mathpar}
Let 
\begin{mathpar}
%J_Q=\displaystyle{\bigcup_{t=0}^nJ_{t}}
%\qquad
\Pred\,'=\Pred\land (\beta_{j_0}=\alpha_Q\land \Pred_Q)
\\
%\forall j\in J_t,\,\gamma_j= \displaystyle{\mybigdotcup_{t=0}^n \gamma_{t j}}\text{\quad and\quad}
\Post\,'=\Post\uplus\Post_Q
\end{mathpar}
Then, we have\\[-2ex]
	\[ P[Q]_{j_0}  
	\models
	{\openrule
		{
			\set{\gamma}\uplus\vis{\beta_j^{j\in J\setminus \{j_0\}}}, 
			\Pred\,',  \Post\,'
			 }
		{\ostate{s_i^{i\in L\uplus L_Q}} \OTWeakarrow {\alpha}
			\ostate{s_i'^{\, i\in L\uplus L_Q}}}
	}
	\]
\end{lemma}
\begin{proof}
By  Lemma \ref{lem-rel-OT-WOT} we can decompose the WOT of $Q$ into  a series of $k+1$ and $k'+1$ tau open transitions and an observable open transition:

\[Q\models{\openrule
	{
		\set{\beta^2},
		 \Pred^2,  
		\Post^2}
	{\ostate{s^2_{0}}\OTarrow {\alpha_Q}
		\ostate{s^2_{1}}}
}
\]

such that 
$s_i^{i\in L_Q}=s^1_{0} \wedge s^1_{(k+1) i}=s^2_{0} \wedge  s^2_{1}= s^3_{0} \wedge s^3_{(k'+1) i}={s'}_i^{i\in L_Q}$\\

$\displaystyle{\set{\gamma}=\mybigdotcup_{h=0}^{k} \vis{\set{\beta^1_h}} \dotcup  \vis{\set{\beta^2}} \dotcup \mybigdotcup_{h=0}^{k'} \vis{\set{\beta^3_h}}}
\\
\Pred_Q=\bigwedge_{h=0}^{k}\!\big(\Pred^1_h\bigotimes_{j=0}^{h-1}\subst{\Post_j^1}\big)\land\Pred^2\bigotimes_{h=0}^{k}\subst{\Post_h^1}\land \bigwedge_{h=0}^{k'}\!\Pred_h^3\big(\bigotimes_{j=0}^{h-1}\subst{\Post^3_j}\subst{\Post^2}  \bigotimes_{h=0}^{k}\subst{\Post_h^1}\big)
\\
\Post_Q=\bigotimes_{h=0}^{k'}\Post^3_h\subst{\Post^2}\bigotimes_{j=0}^{k}\subst{\Post^1_j}
$

\item For the first $k$ open tau transitions, by Definition \ref{def:Non-ObsTau} $P$ can necessarily make a tau open transition if the hole indexed $j_0$ makes a tau action. So by Lemma \ref{lem-compose} we obtain $k$ open transitions in the form: 
\[P[Q]_{j_0}\!\models		
\openrule
    {
       \set{\beta^1_h},{\Pred^1_h},{\Post^1_h}   }
         {\ostate{s^1_h\uplus s^{i\in L}_{i} } \OTarrow {\tau} \ostate{s^1_{h+1}\uplus s^{i\in L}_{i}}}\]

\item For the observable open transition.  By Lemma \ref{lem-compose} with the lemma hypotheses we obtain:\\ 
	\[ P[Q]_{j_0}  
	\models
	{\openrule
		{
			{\beta^i_{p j}}^{(j\in J\setminus\{j_0\})} \uplus \set{\beta_2}, 
			\Pred\land\Pred^2\land \alpha_t=\beta^i_{p j_0 },  
			\Post\uplus \Post^2}
{\ostate{s^{i\in L}_{i}\uplus s_0^2} \OTarrow {\alpha_p} \ostate{{s'}^{i\in L}_{i}\uplus s_1^2}}
	}
	\]

By Lemma \ref{lem-rel-OT-WOT} from cases $(1)$, $(2)$ and $(3)$ we obtain: 
\[P[Q]_{j_0}\!\models		
\openrule
    {
\set{\gamma'}, \Pred_c,
\Post_c}
         {\ostate{s^1_0\uplus s^{i\in L}_{i}} \OTWeakarrow {\alpha^l} 
	\ostate{{{s'}^{i\in L}_{i}\uplus s_{k'}^3}}}\] 
where
\[\set{\gamma'}=\mybigdotcup_{h=0}^k\vis{\set{\beta^1_h}} \dotcup \vis{{\beta^i_{p j}}^{(j\in J\setminus\{j_0\})} \uplus \set{\beta_2}} \dotcup \mybigdotcup_{h=0}^{k'}\vis{\set{\beta^3_h}}\]
\[
\Pred_c=\begin{array}[t]{l}\displaystyle\bigwedge_{h=0}^k{\Pred^1_h}\subst{\bigotimes_{i=0}^{h-1}\Post_i^1}\land
\left(Pred\land\Pred^2\land \alpha_t=\beta^i_{p j_0 }\right)\subst{\bigotimes_{i=0}^{k}{\Post_i^1}}\land\\\displaystyle
\bigwedge_{h=0}^{k'}{\Pred^3_h}\subst{\bigotimes_{i=0}^{h-1}{\Post_i^3}\otimes({\Post\uplus \Post^2})\otimes\bigotimes_{i=0}^{k}{\Post_i^1}}
\end{array}
\]

\[\Post_c=(\bigotimes_{i=0}^{k'}{\Post_i^3})\otimes({\Post\uplus \Post^2})\otimes\bigotimes_{i=0}^{k}{\Post_i^1}
\]
Note that we have $s_t=s^1_{0} \wedge s^1_{(k+1) i}=s^2_{0} \wedge  s^2_{1}= s^3_{0} \wedge s^3_{(k'+1) i}=s_{t+1}$
\end{proof}

\begin{lemma}[Weak open transition composition]\label{lem-Weakcompose} Suppose that we have a weak open automaton such that the WOTs cannot observe silent actions (see Definition \ref{def:Non-ObsTau}).
	Suppose $j_0\in J$ and $\gamma_{j_0}=[\beta_0..\beta_n]$ and additionally:\\[-2ex]
\begin{mathpar}
%\[
P\models{\openrule
	{
		\gamma_j^{j\in J}, 
		\Pred,  
		\Post}
	{\ostate{s_i^{i\in L}} \OTWeakarrow {\alpha}
		\ostate{s_i'^{\, i\in L}}}
}\\
\text{~~and for all $t\in[0..n]$~~}
Q\models{\openrule
	{
		\gamma_{t j}^{j\in J_t},
		 \Pred_t,  
		\Post_t}
	{\ostate{s_{t i}^{i\in L_Q}} \OTWeakarrow {\alpha_t}
		\ostate{s_{(t+1) i}^{\, i\in L_Q}}}
}%\]
\end{mathpar}
Let 
\begin{mathpar}
J_Q=\displaystyle{\bigcup_{t=0}^nJ_{t}}
\qquad
\Pred\,'=\Pred\land \bigwedge_{t=0}^{n}(\alpha_t=\beta_t\land \Pred_t)\bigotimes_{i=0}^{t-1}\Post_i
\\
\forall j\in J_t,\,\gamma_j= \displaystyle{\mybigdotcup_{t=0}^n \gamma_{t j}}\text{\quad and\quad}
\Post\,'=\Post\uplus\bigotimes_{t=0}^{n}
		\Post_t
\end{mathpar}
Then, we have\\[-2ex]
	\[ P[Q]_{j_0}  
	\models
	{\openrule
		{
			\gamma_j^{j\in (J\setminus\{j_0\}) \uplus J_Q}, 
			\Pred\,',  \Post\,'
			 }
		{\ostate{s_i^{i\in L\uplus L_Q}} \OTWeakarrow {\alpha}
			\ostate{s_i'^{\, i\in L\uplus L_Q}}}
	}
	\]
\end{lemma}
\TODO{Proof starts from here}

@@

Note that, for $l\in\{1,2,3\}$ if $(\beta^l_p)_{j_0}=\tau$ because of Definition \ref{def:Non-ObsTau}, $P$ necessarily makes a $\tau$ open transition and remains in the same state, ($s_p = s_{p+1}$). Thus without loss of generality, we can bypass such an open transition and obtain another decomposition of the WOT without the open transition that requires ${\beta^l_{p{j_0}}}=\tau$. We can thus  suppose that for all $p$ and $l$ we have ${\beta^l_{p{j_0}}}\neq\tau$.

Moreover, we have for all $t\in[0..n]$:
\begin{mathpar}
Q\models{\openrule
	{
		\gamma_{t j}^{j\in J_t},
		 \Pred_t,  
		\Post_t}
	{\ostate{s_{t i}^{i\in L_Q}} \OTWeakarrow {\alpha_t}
		\ostate{s_{(t+1) i}^{\, i\in L_Q}}}
}		
\end{mathpar}
%In the same way,  by Lemma \ref{lem-rel-OT-WOT} we can decompose $\WOT_p$ into  a series of $m'+n'$ tau open transitions and a single observable open transition:
%\[Q\models{\openrule
%			{
%				(\beta^2_j)^{j\in J\cap\Holes(Q)}, {{\Pred'^{2}}},  
%				{{\Post'^{2}}}}
%				{\ostate{s_{p i}^{i\in L_Q}} \OTarrow {\alpha_p}
%				\ostate{s_{(p+1) i}^{i\in L_Q}}}
%				}%\]
%\]

We proceed by iteration on the list of  OTs of $P$ identified above, these OTs are of the form, for $l\in\{1,2,3\}$: 
\[P\!\models		
\openrule
    {
       {(\beta^l_{p j})^{j\in J}},\Pred_p,\Post_p   }
         {\ostate{s^{i\in L}_{pi}} \OTarrow {\alpha^l} \ostate{s^{i\in L}_{(p+1)i}}}\]
with $\alpha^2=\alpha$ and $\alpha^1=\alpha^3=\tau$

@@

By Lemma \ref{lem-rel-OT-WOT} from cases $(1)$, $2$ and $(3)$ we obtain: 
\[P[Q]_{j_0}\!\models		
\openrule
    {
\set{\gamma}, \Pred_c,
\Post_c}
         {\ostate{s^1_0\uplus s^{i\in L}_{pi}} \OTWeakarrow {\alpha^l} 
	\ostate{{s^{i\in L}_{(p+1)i}\uplus s_{k'}^3}}}\] 
where
\[\set\gamma=\mybigdotcup_{h=0}^k\vis{\set{\beta^1_h}} \dotcup \vis{{\beta^i_{p j}}^{(j\in J\setminus\{j_0\})} \uplus \set{\beta_2}} \dotcup \mybigdotcup_{h=0}^{k'}\vis{\set{\beta^3_h}}\]
\[
\Pred_c=\begin{array}[t]{l}\displaystyle\bigwedge_{h=0}^k{\Pred^1_h}\subst{\bigotimes_{i=0}^{h-1}\Post_i^1}\land
\left(Pred_p\land\Pred^2\land \alpha_t=\beta^i_{p j_0 }\right)\subst{\bigotimes_{i=0}^{k}{\Post_i^1}}\land\\\displaystyle
\bigwedge_{h=0}^{k'}{\Pred^3_h}\subst{\bigotimes_{i=0}^{h-1}{\Post_i^3}\otimes({\Post_p\uplus \Post^2})\otimes\bigotimes_{i=0}^{k}{\Post_i^1}}
\end{array}
\]

\[\Post_c=(\bigotimes_{i=0}^{k'}{\Post_i^3})\otimes({\Post_p\uplus \Post^2})\otimes\bigotimes_{i=0}^{k}{\Post_i^1}
\]
Note that we have $s_t=s^1_{0} \wedge s^1_{(k+1) i}=s^2_{0} \wedge  s^2_{1}= s^3_{0} \wedge s^3_{(k'+1) i}=s_{t+1}$

 \item Else, $j_0\not\in J$. $t$ does not move  and $Q$ does not move, i.e. we have: $\ostate{(s_{t})} = \ostate{({s}_{(t+1)})}$
By lemma~\ref{lem-compose-2} we have:
\[P[Q]_{j_0}\!\models		
\openrule
    {
       {(\beta^l_{p j})^{j\in J}},\Pred_p,\Post_p   }
         {\ostate{s^{i\in L}_{pi}\uplus s_t} \OTarrow {\alpha^l} \ostate{s^{i\in L}_{(p+1)i}\uplus s_t}}
\]
and thus
\[P[Q]_{j_0}\!\models		
\openrule
    {
       \vis{{(\beta^l_{p j})^{j\in J}}},\Pred_p,\Post_p   }
         {\ostate{s^{i\in L}_{pi}\uplus s_t} \OTWeakarrow {\alpha^l} \ostate{s^{i\in L}_{(p+1)i}\uplus s_t}}
\]

\end{itemize}
